---
title: "244 Lab 2"
author: "Molly Williams"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load}

library(tidyverse)
library(janitor)
library(MASS)
library(RColorBrewer)
library(factoextra)
library(shinythemes)
library(ggrepel)


# To install ggbiplot, install development version
library(devtools)
install_github("vqv/ggbiplot")
library(ggbiplot)
```



If you want to override (set) a function maually throughout an entire workspace, you can assign the function name explicitly:

```{r}

select <- dplyr::select #general override, tells R to use the select function from the dplyr package
```

###1. Ordinal logistic regression for political party affiliation by ethnicity, age, sex

```{r}

pol_party <- read_csv("pol_party.csv") %>% 
  clean_names() %>% 
  select(participant_id:read_carefully)

# a little more wrangling:

pol_df <- pol_party %>% 
  select(birth_year, gender, ethnicity, political_party) %>% 
  filter(birth_year > 1900, political_party != "NA", political_party <= 7) %>% 
  mutate(age = 2016 - birth_year)


# Political party: 1 = very liberal, 7 = very conservative

# When doing OLR we want our dependent variable to be of class ORDERED FACTOR

pol_df$political_party <- factor(pol_df$political_party, ordered = TRUE, levels = c("1", "2", "3", "4", "5", "6", "7"))


# Update our categorical predictor variables to factors too:

pol_df$ethnicity <- factor(pol_df$ethnicity)
pol_df$gender <- factor(pol_df$gender)

```


#### Basic data exploration and visualization

```{r}

# Create new df with counts by race: - but whites are so heavily represented in responses that it's not useful to analyze based on this factor
counts_eth <- pol_df %>% 
  group_by(ethnicity, political_party) %>% 
  tally()

# Create new df with counts by gender:
counts_g <- pol_df %>% 
  group_by(gender, political_party) %>% 
  tally()

# just because you don't get an error message doesn't mean your method is good (:


ggplot(pol_df, aes(x = political_party)) +
  geom_histogram(aes(fill = ethnicity), stat = "count") +
  scale_fill_brewer(palette = "YlGnBu") +
  facet_wrap(~ethnicity, scales = "free")

```


#### THESE DATA ARE PRECARIOUS AND N IS TOO SMALL!!! FOR GROUPS
```{r}

### This is not publishable data!! Data is so biased that it gives the audience an idea that something is happening, but the sample size is super small

pol_model <- polr(political_party ~ age + gender + ethnicity, data = pol_df)

summary(pol_model)

exp(coef(pol_model))

# Once we've decided on a model, we can always make predictions using it:

predictions <- predict(pol_model, type = "probs")

df <- data.frame(pol_df, predictions) # generates a model prediction for each point, can visualize in ggplot and see how it aligns with actual selections made by respondents

```


### 2. Principal components analysis (PCA)
#### ...Using World Bank environmental indicators


```{r}
 wb_10 <- read_csv("wb_10.csv") %>% 
  select(-X1) %>% # removes this column 
  column_to_rownames('code') # changes all of the row labels to country codes

wb_pca <- prcomp(wb_10[3:8], scale = TRUE) # data are on different scales, so perform scale function

summary(wb_pca)

ggbiplot(wb_pca) +
  geom_text_repel(label = rownames(wb_10), size = 2, segment.size = 0.2) +
  theme_bw()

# factoextra package (built to help visualize outcomes of ordination methods)

# scree plot:

fviz_screeplot(wb_pca)

# See contributions of variables to different components:

fviz_contrib(wb_pca, choice = "var", axes = 1, top = 6) # can change axes to 2 to look at PC2

# Biplot showing points and vectors:

fviz_pca_biplot(wb_pca,
                col.var = "cos2", 
                label = "var",
                repel = "TRUE")

# col.var = cos2 is a metric of how well represented variables are in PC1 (highly related with what is on pc2, but can check the loadings for where variables start having the greatest impact)
  
```


### 3. SHINY APP TIME!










